Currently, services requirements are defined in a contract, but there is no standardized method for transforming contractual requirements into a mathematical model (requirements model) that a computer can use for verification of service delivery. This requirements model forms the base-line against which the delivered service is benchmarked. The model incorporates both the ideal and acceptable service provision. By analysis of the TSO services defined in \cite{EnerginetAncillary}, the potential DSO services defined in \cite{ding2013development} and asset management services, we have developed a method for modeling the service delivery over time. 
%The method presented in this section outlines the steps to transforming contractual requirements into a requirements model.
The method consists of the following six steps:
%form have been identified as a generic method for modeling ancillary services. The steps are exemplified by DSO services defined in  and TSO services defined in :
\begin{enumerate}
  \item Identify physical parameters defining the service.
%  \begin{itemize}
%    \item For example: Power production or consumption, measured grid frequency, time measurements. Including maximum measuring sensitivities.
%  \end{itemize}
  \item Identify the dynamic behaviors of the service related to system parameters (if any).
%  \begin{itemize}
%    \item Example: Primary frequency regulation comes with a linear behavior between grid frequency and generator set-point. Power-cap has a dynamic relationship between feeder load and the controllable load power in order to keep the total feeder load at a $P_{DSO,Ref}$ value.
%    \item PowerMax is not dynamic. The aggregator must control $\Delta P_{Agg}$ to ensure that he does not violate $P_{max,Agg}$. But the service does not require a dynamic behavior related to a system parameter like primary frequency regulation and PowerCap.
%  \end{itemize}
  \item Identify the physical size of the service and the tolerated error. % Both ideal service and minimum required service.
%  \begin{itemize}
%    \item Physical size is for example $P_{max,Agg}$ for PowerMax service or regulation bid for primary frequency regulation.
%    \item Tolerance is for example $P_{max,Agg}+P_{tolerance}$ for the PowerMax service and allowed dead-band for DK1 primary reserve.
%  \end{itemize}
  \item Identify the ideal response time of the service and acceptable response.
%  \begin{itemize}
%    \item Most contracts comes with some timing specifying how fast the service provider must act.
%    \item E.g. DK1 primary reserve must provide 50 \% of service within 15 s and 100 \% within 30 s.
%    \item The ideal service is for example an instantaneous step in power to 100 \% of set-point for DK1 primary reserve.
%  \end{itemize}
  \item Based on the dynamics, size and timing of the service, as well as the tolerated errors from points 1--4, develop a time series for ideal and acceptable service provision. The model will be a set of time series: $\mathbf{x}_{ideal}(t)$ for ideal response and $\mathbf{x}_{acc}(t)$ for acceptable response. Both time series can be a scalar or a vector, e.g. $\mathbf{x}_{acc}(t)$ can be formed by a set of upper and lower tolerance bounds or simply by an upper bound.
  %ยง$\mathbf{x}_{ideal}(t)$ and $\mathbf{x}_{acc}(t)$ can be a pair of values, e.g. minimum and maximum tolerance limits, for some services and may be only a single value, e.g. minimum or maximum tolerance, for others. 
  \item Identify how the service error is to be measured.
\end{enumerate}

We identify three different types of service: reference tracking, band service or a maximum/minimum cap. The error measure, $e(t) \in \mathbb{R}$, for each of the service types is defined in the following subsections. This approach was initially introduced in \cite{bondy2014performance}, and is further refined in this work. 

%The error, $e(t) \in \mathbb{R}$, in the delivery of a service can be calculated using a reference tracking, band service or a maximum/minimum cap approach. The proposed approach depends on the specific service and contract.\bondynote{These last two paragraphs can probably be merged}


\subsection*{Reference tracking}
Reference tracking error can be calculated as:
\begin{equation}\label{eq:ref_error}
e(t) = x_{meas}(t) - x_{ideal}(t).
\end{equation}
This definition will lead $e<0$ for measured values below the ideal and $e>0$ for values above the ideal. In this case $\mathbf{x}_{acc}(t)$ will be a band around $x_{ideal}(t)$, and the values of $\mathbf{x}_{acc}(t)$ do not need to be symmetric.

\subsection*{Band service}
The ideal response in a band service is defined as $ \mathbf{x}_{ideal}(t)= [x_{min}(t),x_{max}(t)]$. The error in the band service can therefore be estimated by:
\begin{equation}\label{eq:band_error}
e(t)=
\begin{cases}
x_{meas}(t) - x_{min}(t) , & x_{meas}(t) < x_{min}(t)  \\
0, & x_{min}(t) \leq x_{meas}(t) \leq x_{max}(t) \\
x_{meas}(t) - x_{max}(t), & x_{meas}(t)  > x_{max}(t).  
\end{cases}
\end{equation}
In this case, the $\mathbf{x}_{acc}(t)$ is a set of values that surrounds the band defined by $ \mathbf{x}_{ideal}(t)$, as seen in Fig.~\ref{fig:RefErr}. The values of $\mathbf{x}_{acc}(t)$ do not need to be symmetric around the band.

\subsection*{Cap service}
%Maximum/minimum cap error only counts the error when performance is above/below some ideal value. 
In cap services, error is only tracked when $x_{meas}(t)$ is either above or below a given a limit value.
Maximum cap error is calculated as shown in \eqref{eq:maxmin_cap} and minimum cap can be similarly calculated. In \eqref{eq:maxmin_cap}, $x_{max}(t)$ is the ideal maximum limit according to the service contract:

\begin{equation}\label{eq:maxmin_cap}
e(t)=
\begin{cases}
x_{meas}(t)-x_{max}(t), & x_{meas}(t) > x_{max}(t) \\
0, & x_{meas}(t) \leq x_{max}(t).
\end{cases}
\end{equation}
In the cap service, $x_{acc}(t)$ is a limit that either lies below $x_{min}(t)$ or above $x_{max}(t)$.

Having defined the method for formulating the service models, we will show how these can be used for performance assessment of services.
