% TEX root = ../Thesis.tex
\chapter{List of Publications}
\section*{Papers included in this thesis}
The following papers can be found in the appendices in the order the concepts are presented in the thesis. Here they are listed chronologically.
\subsection*{Papers presented at peer-reviewed conferences}
\begin{itemize}
	\item[D)] D.E.M. Bondy, G.T. Costanzo, K. Heussen, and H.W. Bindner. ``Performance assessment of aggregation control services for demand response''. In: \emph{Innovative Smart Grid Technologies Conference Europe (ISGT-Europe), 2014 IEEE PES}. Istanbul, Turkey: IEEE, Oct. 2014, pp. 1–6. \textsc{Doi}: 10.1109/ISGTEurope.2014.7028779.
	\item[A)] D.E.M. Bondy, K. Heussen, O. Gehrke, and A. Thavlov.``A Functional Reference Architecture for Aggregators''. eng. In: \emph{Proceedings of the 20th IEEE International Conference on Emerging Technologies and Factory Automation (ETFA 2015)}. Luxembourg, Luxembourg: IEEE, Sep. 2015. \textsc{Doi}: 10.1109/ETFA.2015.7301638
	\item[B)] D.E.M. Bondy, O. Gehrke, A. Thavlov, K. Heussen, A.M. Kosek, and H.W. Bindner. ``Procedure for Validation of Aggregators Providing Demand Response''. \emph{To be presented at Power Systems Computational Conference (PSCC 2016)}. Genoa, Italy. Jun. 2016.
\end{itemize}

\subsection*{Paper submitted to peer-reviewed journal}
\begin{itemize}
	\item[E)] D.E.M Bondy, A. Thavlov, and J.B.M. Tougaard, Kai Heussen. ``Method for Ancillary Service Modeling and Performance Assessment''. \emph{Submitted to IEEE Transactions on Smart Grids}.
\end{itemize}

\subsection*{Draft journal paper}
\begin{itemize}
\item[C)] D.E.M. Bondy, J. MacDonald, E. Kara, O. Gehrke, K. Heussen, S. Kiliccote and H.W. Bindner. ``Redefining Requirements of Ancillary Services for Technology Agnostic
	Sources''. \emph{Draft.}
\end{itemize}

\section*{Other publications not included in this thesis}
The following publications have been prepared during the course of the PhD study. They are omitted from this thesis since they are not directly related to the primary objective of the PhD project, or the relevant contributions were subsequently published in the papers mentioned above.
\begin{itemize}
	\item D.E.M. Bondy, G. Tarnowski, K. Heussen, and L.H. Hansen. ``Operational Scenario: Manual Regulating Power''. \emph{Tecnical report for the iPower Consortium}. 2013.
	\item D.E.M. Bondy, and A. Thavlov. ``FLECH PowerMax Service Requirement Specification''. \emph{Technical report for the iPower Consortium}. 2014.
\end{itemize}

\subsection*{Papers as contributing author}
\begin{itemize}
	\item G.T. Costanzo, O. Gehrke, D.E.M. Bondy, F. Sossan, H.W. Bindner, J. Parvizi, and H. Madsen. ``A Coordination Scheme for Distributed Model Predictive Control: Integration of Flexible DERs''. In: \emph{Proc. 4th IEEE PES Int. Conf. Innovative Smart Grid Technologies Europe}. Copenhagen, Denmark: IEEE, Oct. 2013. \textsc{Doi}: 10.1109/ISGTEurope.2013.6695474.
	\item K. Heussen, D.E.M. Bondy, J. Hu, O. Gehrke, L.H. Hansen. ``A Clearinghouse Concept for Distribution-Level Flexibility Services''. In: \emph{Proc. 4th IEEE PES Int. Conf. Innovative Smart Grid Technologies Europe}. Copenhagen, Denmark: IEEE, Oct. 2013. \textsc{Doi}: 10.1109/ISGTEurope.2013.6695483.
	\item X. Han, A.M. Kosek, D.E.M. Bondy, H.W. Bindner, S. You, D.V. Tackie, J. Mehmedalic, and F. Thordarson. ``Assessment of Distribution Grid Voltage Control Strategies in View of Deployment''. In: \emph{Intelligent Energy Systems (IWIES), 2014 IEEE International Workshop on}. IEEE. 2014, pp. 46–51. \textsc{Doi}: 10.1109/IWIES.2014.6957045.
\end{itemize}
